---
title: "R Programming Presentation"
author: "Daniel Dean, Jessica Nunez, Erin Wall, Chayou Zhai"
date: "12/5/2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction:

Jeopardy has been running for 35 seasons. In this project our task was to create a question, and later use R Programming to manipulate raw show data to answer our question. Our raw data included questions, answers, daily double, air date, and notes. We were excited to get more experience in mapping in R, and this seemed like a good opportunity.

## Research Questions:
Which countries are most frequently-mentioned in Jeopardy, and how does that relate to the frequency of daily doubles? 

Is there a correlation between land area and mentions in Jeopardy?

Is there a correlation between land area or GDP and mentions in Jeopardy?

## Methods:

### Data Processing

- We started with Github user `jwolle`'s extensive Jeopardy archives, which span from September 1984 to July of this year. These were easy to read in using `tidyverse` tools

- To generate a list of country names, we drew on a set up country synonyms from the `rworldmap` package, as well as an HTML table of country names, adjectives, and demonyms from Wikipedia using the `rvest` web scraping package. 

- We used regex matching with package `stringr` to extract instances of names in our country list from Jeopardy questions and answers.

##

- United Nations ISO3 codes were the backbone of our dataset, allowing us to associate terms (e.g. we linked "China","Chinese","PRC",  and "People's Republic of China", to the ISO3 code "CHN")

- Naturally, there were some complications; most notably, "Indian" frequently referred to Native Americans. We were able to partially mitigate this with Category context (e.g. "Colonial America", "Native Americans"). 

- Another interesting wrinkle was words like "French" and "Dutch", which are both adjectives and demonyms, being double-counted, briefly making France appear to be the most popular country in Jeopardy history. 

### Mapping

- We matched our ISO3-linked jeopardy data to geographic data (vector polygons defining borders, among other information) from the `rworldmap` package using the `join` family of functions from `dplyr`

- Once summarized (to optimize memory usage) and converted into an `sf`, or *simple feature* format, we were able to pass these objects to theinteractive mapping package `leaflet`, among other downstream applications.


### Statistics

- The resulting table, with the original jeopardy data, our fequency tallies, and the ISO3 codes, was matched to a world map <*~`LowRest~` or something*> bundled with `rworldmap`, which also included ISO3 codes. 

```{r, }


```

- We used relatively basic analyses in this project

- For the question of which countries were most frequently mentioned, we simply grouped by iso3 codes, and counted the number of instances of each, either over all time, or within a given year.

- To explore the question of how reference frequency is realted to land area and GDP, we fit linear models to test the correlation of land area to mentions.

 
## Results:

### Frequency

- Unsurprisingly, America is the most-commonly mentioned country, at 16,395 references over the years, well ahead of runner-up France with 6,847, and the UK at 3,844.

- The US, India, and Mexico were the only non-European countries that made the all-time top 10 (also those with >2,000 references), although Japan and Canada were relatively close behind in the 1,800s.  

- While additional false positives undoubtedly made it through, India was genuinely popular in geography and history questions, as well as the subject of some tailor-made categories ("India", "A Passage to India", "Punjab", etc.)


```{r Leaflet Map, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rworldmap)
library(sf)
library(leaflet)
library("dplyr")
library(viridisLite)
library(janitor)
library(lubridate)
library(broom)
library(plotly)

country_data_all <-read_csv("country_all_iso_all.csv", )

countriesLow <- countriesLow %>%
  st_as_sf

#Temporarily removing air date (not sure how to animate/facet/etc. in Leaflet)

country_geom_full<- country_data_all %>%
  left_join(countryExData, by = c("iso3" = "ISO3V10")) %>%
  group_by(iso3) %>%
  mutate(mean_value = mean(value)) %>%
  add_tally(name = "count") %>%
  ungroup() %>%
  select(country, count, mean_value, iso3) %>%
  distinct() %>%
  mutate(iso3 = toupper(iso3)) %>%
  rename(ISO3 = iso3)

country_geom_map_data <- country_geom_full %>%
  mutate(ISO3 = as.factor(ISO3)) %>%
  dplyr::full_join(countriesLow) %>%
  clean_names() %>%
  st_as_sf

pal <- colorQuantile(
  palette = "Greens",
  domain = country_geom_map_data$count)

popup_info<- paste0("<b>Country:</b> ",
                    country_geom_map_data$name, "<br/>",
                    "<b>Population:</b>",
                    country_geom_map_data$pop_est, "<br/>",
                    "<b>Count:</b>",
                    country_geom_map_data$count, "<br/>",
                    "<b>Mean Value:</b> ",
                    round(country_geom_map_data$mean_value))




leaflet(country_geom_map_data) %>%
  addTiles() %>%
  addPolygons(color = ~pal(count), popup = popup_info) %>%
  addLegend(pal = pal, values = ~count, position = 'topright')


```


### Is there a correlation between land area and mentions in Jeopardy?


## Land Area and Frequency:

- We'd speculated that the popularity of relatively small European countries and (to a lesser extent) Carribean Islands could lead to a relatively weak, or even negative, correlation between country size and popularity in questions/answers, but found a fairly consistent, positive correlation.

- The country with biger land area or higher GDP has higher chance to be mentioned in Jeopardy.

- 

- Russia, being both relatively popular and a massive 16,679,993 km^2, almost single-handedly sets this trend (see animation), although the next-largest countries in Canada (9458907 km^2), the USA (9210753 km^2) and China (9198094 km^2) are relatively popular as well. 


jeopardy_all_season  <-  country_data_all%>%
  mutate(year = year(air_date)) %>%
  group_by(iso3, year) %>%
  add_tally(name = "count") %>%
  ungroup() %>%
  mutate(iso3 = toupper(iso3),
         iso3 = as.factor(iso3)) %>%
  rename(ISO3V10 = iso3) %>%
  full_join(countryExData) %>%
  filter(!is.na(count)) %>%
  rename(iso3 = "ISO3V10") %>%
  clean_names() %>%
  mutate(ratio = count / landarea)



ggplotly(jeopardy_all_season %>%
  select(iso3, landarea, year, epi_regions, count) %>%
  #filter(iso3 != 'USA') %>%
  distinct() %>%
  group_by(year, epi_regions) %>% #summarize(number = n()) %>% arrange(desc(number)) %>%
  drop_na() %>%
  nest() %>%
  mutate(model = purrr::map(data, ~lm(count ~ landarea, data = .x) %>% 
                              tidy() %>% select(term, estimate))) %>%
  unnest(c(model)) %>% #arrange(desc(p.value)) <-- all p values are <0.05
  pivot_wider(names_from = term, values_from = estimate) %>%
  clean_names %>%
  rename(slope = landarea) %>%
  unnest(data) %>%
  ggplot() +
  geom_point(aes(x = landarea, y = count, frame = year, color = epi_regions, label = iso3)) +
  geom_abline(aes(slope = slope, intercept = intercept, frame = year, color = epi_regions))
)

```

## GDP and Frequency:

- GDP also showed a positive correlation with frequency of mentions; in this case the USA played an outsize role


*
ggplotly(jeopardy_all_season %>%
           select(iso3, gdp_capita_mrya, year, epi_regions, count) %>%
           #filter(iso3 != 'USA') %>%
           distinct() %>%
           group_by(year) %>% #summarize(number = n()) %>% arrange(desc(number)) %>%
           drop_na() %>%
           nest() %>%
           mutate(model = purrr::map(data, ~lm(count ~ gdp_capita_mrya, data = .x) %>% 
                                       tidy() %>% select(term, estimate))) %>%
           unnest(c(model)) %>% #arrange(desc(p.value)) <-- all p values are <0.05
           pivot_wider(names_from = term, values_from = estimate) %>%
           clean_names %>%
           rename(slope = gdp_capita_mrya) %>%
           unnest(data) %>%
           ggplot() +
           geom_point(aes(x = gdp_capita_mrya, y = count, frame = year, color = epi_regions, label = iso3)) +
           geom_abline(aes(slope = slope, intercept = intercept, frame = year))
)

```


## Daily Doubles

### *Placeholder: for some reason my R session keeps ~freezing (the last merge step uses a lot of memory eve after I stop) when I try running `daily_double_and_frequency`*

```{r}
frequency_map_test <- country_all_iso_all %>%
  select(value, daily_double, air_date, country, iso3) %>%
  distinct() %>%
  mutate(year = ymd(air_date)) %>%
  group_by(iso3, year) %>%
  add_tally(name = "season_count") %>%
  ungroup() %>%
  mutate(daily_double = fct_recode(daily_double, "TRUE" ="yes", "FALSE" = "no"),
         daily_double = as.logical(daily_double)) %>%
  group_by(iso3) %>%
  summarize(count = n(), mean_value = mean(value), daily_double_count = sum(daily_double)) %>%
  full_join(countriesLow_sf, by = c("iso3" = "ISO3")) %>%
  st_as_sf






ggplotly(ggplot() +
  
  
  
  geom_sf(data = frequency_map_test, aes(fill = daily_double_count, color = mean_value)) +
  
  
  
  scale_fill_viridis_c() +
    scale_color_viridis_c(option = 'plasma')
)

```


## Conclusion

## Tutorial:
### Data Manipulation

- Read in Jeopardy question data compiled by Github user `jwolle1`; initially (at all?) used Season 1

- Used `read_csv` from the `readr` package, although we manully downloaded the full dataset, a zipped file.

- The raw data already conformed to tidy data conventions, so no special pre-processing was needed on this front.

## Next Steps . . .
- We needed a list of names associated with countries

- Our basis was the `CountrySynomnyms` dataframe from the `rworldmap` package, this included up to 8 synonymous names for every country recognized as of 2005 (as well as historical country names), along with 3-letter abbreviations following the ISO3 standard. 

- We used the `pivot_longer` function from `tidyr` to convert this dataframe to two columns: ISO3 names and names (NAs were removed).

## Next Steps Continued . . .
- To expand this dataset, names from the "lengthend" country names dataframe were matched against a list of country adjectives and demonyms (e.g. "Russian", "Russians") scraped from the Wikipedia page using the `rvest` package.

- These additional names were also converted to a single column, and matched to ISO3 codes.

## Matching Names
- We then used the `str_detect` and `str_extract_all` functions from `stringr` in tandem to locate and extract matches in Jeopardy questions or answers. 
  
- We avoided some false positives (e.g. "Indiana" includes the string "India") by excluding any match that was follwed by a letter (our target list included both singualr and plural forms of country adjectives/demonyms). 
  
## Finally . . .

* Because `str_extract_all` generates a list, we used <an option> to add NA values to equialize the products' lengths, and `unnest` from <`?`> to convert these into separate rows.

* We added the source (question or answer) as a metadata column, and merged both derived datasets to get a total frequency.


## Mapping:

- The resulting table, with the original jeopardy data, our fequency tallies, and the ISO3 codes, was matched to a world map <*~`LowRest~` or something*> bundled with `rworldmap`, which also included ISO3 codes.

```{r}

```

### Leaflet Package
```{r echo= TRUE, tidy=TRUE, message=FALSE, warning= FALSE}
library(tidyverse)
library(rworldmap)
library(sf)
library(leaflet)
library("dplyr")
library(viridisLite)
library(janitor)

  country_data_all <-read_csv("country_all_iso_all.csv", )

countriesLow <- countriesLow %>%
  st_as_sf

#Temporarily removing air date (not sure how to animate/facet/etc. in Leaflet)

country_geom_full<- country_data_all %>%
  left_join(countryExData, by = c("iso3" = "ISO3V10")) %>%
  group_by(iso3) %>%
  mutate(mean_value = mean(value)) %>%
  add_tally(name = "count") %>%
  ungroup() %>%
  select(country, count, mean_value, iso3) %>%
  distinct() %>%
  mutate(iso3 = toupper(iso3)) %>%
  rename(ISO3 = iso3)

country_geom_map_data <- country_geom_full %>%
  mutate(ISO3 = as.factor(ISO3)) %>%
  dplyr::full_join(countriesLow) %>%
  clean_names() %>%
  st_as_sf

pal <- colorNumeric(
  palette = "Greens",
  domain = country_geom_map_data$count)

popup_info<- paste0("<b>Country:</b> ",
                    country_geom_map_data$name, "<br/>",
                    "<b>Population:</b>",
                    country_geom_map_data$pop_est, "<br/>",
                    "<b>Count:</b>",
                    country_geom_map_data$count, "<br/>",
                    "<b>Mean Value:</b> ",
                    round(country_geom_map_data$mean_value))




leaflet(country_geom_map_data) %>%
  addTiles() %>%
  addPolygons(color = ~pal(count), popup = popup_info)


```
### Plotly
```{r}

```

## Lessons Learned
- Data Version Control
- Importance of planning and communication



