---
title: "R Programming Presentation"
author: "Daniel Dean, Jessica Nunez, Erin Wall, Chayou Zhai"
date: "12/5/2019"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Research Questions
What is the most frequently mentioned country in Jeopardy? ...
## Introduction

## Methods
- Read in Jeopardy question data compiled by Github user `jwolle1`; initially (at all?) used Season 1

- Used `read_csv` from the `readr` package, although we manully downloaded the full dataset, a zipped file.

- The raw data already conformed to tidy data conventions, so no special pre-processing was needed on this front.

## Next Steps . . .
- We needed a list of names associated with countries

- Our basis was the `CountrySynomnyms` dataframe from the `rworldmap` package, this included up to 8 synonymous names for every country recognized as of 2005 (as well as historical country names), along with 3-letter abbreviations following the ISO3 standard. 

- We used the `pivot_longer` function from `tidyr` to convert this dataframe to two columns: ISO3 names and names (NAs were removed).

- To expand this dataset, names from the "lengthend" country names dataframe were matched against a list of country adjectives and demonyms (e.g. "Russian", "Russians") scraped from the Wikipedia page using the `rvest` package.

- These additional names were also converted to a single column, and matched to ISO3 codes.
## Matching Names
- We then used the `str_detect` and `str_extract_all` functions from `stringr` in tandem to locate and extract matches in Jeopardy questions or answers. 
  
- We avoided some false positives (e.g. "Indiana" includes the string "India") by excluding any match that was follwed by a letter (our target list included both singualr and plural forms of country adjectives/demonyms). 
  
## Finally . . .

* Because `str_extract_all` generates a list, we used <an option> to add NA values to equialize the products' lengths, and `unnest` from <`?`> to convert these into separate rows.

* We added the source (question or answer) as a metadata column, and merged both derived datasets to get a total frequency.


## Mapping:

- The resulting table, with the original jeopardy data, our fequency tallies, and the ISO3 codes, was matched to a world map <*~`LowRest~` or something*> bundled with `rworldmap`, which also included ISO3 codes. 

- From there, were were able to use the `leaflet` package to make an interactive world mapwith frequncy of references in jeopardy color-coded with the `scheme we end up using` in <`ggplot?`>

## Leaflet
- We used leaflet package to map out out the countries mentioned, how many times they were mentiones and the airdates when they were mentioned. 

```{r pressure}
library("ggplot2")
library("readr")
library(leaflet)
library(tidyverse)
source("country_dataframes.R")
#To get country_all_isoobject; can comment this pout if already loaded

test_geom <- countriesLow %>% st_as_sf

test_geom_full <- country_all_iso %>%
  group_by(iso3) %>%
  add_tally(name = "count") %>%
  ungroup() %>%
  mutate(iso3 = toupper(iso3),
         iso3 = as.factor(iso3)) %>%
  rename(ISO3 = iso3) %>%
  full_join(test_geom) %>%
  mutate(ISO3 = as.factor(ISO3)) %>%
  clean_names() %>%
  st_as_sf

library("viridisLite")
pal <- colorNumeric(
  palette =  "Greens",
  domain = test_geom_full$count)

popup_info<- paste0("<b>Country:</b> ",
                      test_geom_full$name, "<br/>",
                    "<b>Population:</b> ",
                      test_geom_full$pop_est, "<br/>",
                    "<b>Count:</b> ",
                      test_geom_full$count, "<br/>",
                    "<b>Air Date:</b> ",
                      test_geom_full$air_date)
                    
                    

leaflet(test_geom_full) %>%
  addTiles() %>%
  addPolygons(color = ~pal(count), popup = popup_info)

```


## Results

## Conclusion

## Tutorial

## Lessons Learned






## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
library("ggplot2")
library(leaflet)
library(tidyverse)
source("country_dataframes.R") #To get country_all_isoobject; can comment this pout if already loaded

test_geom <- countriesLow %>% st_as_sf

test_geom_full <- country_all_iso %>%
  group_by(iso3) %>%
  add_tally(name = "count") %>%
  ungroup() %>%
  mutate(iso3 = toupper(iso3),
         iso3 = as.factor(iso3)) %>%
  rename(ISO3 = iso3) %>%
  full_join(test_geom) %>%
  mutate(ISO3 = as.factor(ISO3)) %>%
  clean_names() %>%
  st_as_sf

library("viridisLite")
pal <- colorNumeric(
  palette =  "Greens",
  domain = test_geom_full$count)

popup_info<- paste0("<b>Country:</b> ",
                      test_geom_full$name, "<br/>",
                    "<b>Population:</b> ",
                      test_geom_full$pop_est, "<br/>",
                    "<b>Count:</b> ",
                      test_geom_full$count, "<br/>",
                    "<b>Air Date:</b> ",
                      test_geom_full$air_date)
                    
                    

leaflet(test_geom_full) %>%
  addTiles() %>%
  addPolygons(color = ~pal(count), popup = popup_info)

```

